Blockchain Consensus Protocol with Horizontal Scalability

### 2017-07-26 presentation/presentation.tex
Update presentation

\usepackage{graphicx}
\graphicspath{{../figures/}}

\title{Presentation Title}

% A subtitle is optional and this may be deleted
\subtitle{Optional Subtitle}

@@

% \date{Conference Name, 2013}
\date{2017}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

@@

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
\AtBeginSubsection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

@@

% Section and subsections will appear in the presentation overview
% and table of contents.
\section{Background}

\subsection{TrustChain}

\begin{frame}{TrustChain}{}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-good}
  \centering
  \caption{Every node has a chain. \emph{TX block} is a six-tuple: $t_{i,j} =
    (\texttt{h}(b_{i,j-1}), h_s, h_r, s_s, s_r, m)$, one transaction results in
    two TX blocks---\emph{a pair}.}
  \end{figure}

\end{frame}

\begin{frame}{TrustChain}{}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-bad}
  \centering
  \caption{Fork is two correctly signed TX blocks that belong to the same sender
    and has the same hash pointer but involve different receivers. Only one TX
    block may be in consensus.}
  \end{figure}

\end{frame}

\begin{frame}{TrustChain}
  \begin{itemize}
    \item Everyone has their own chain
    \item Transactions are on arbitrary data $m$
    % \item Transactions make the chains intertwined
    \item Transactions are irrefutable due to hash pointers
    \item No consensus (my thesis)
  \end{itemize}
\end{frame}

\section{My Thesis}
\subsection{TrustChain with Checkpoints}
\begin{frame}{TrustChain with Checkpoints}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-bad-cp}
  \centering
  \caption{CP block is a six-tuple: $c_{i,j} = (\texttt{h}(b_{i,j-1}),
    \texttt{h}(\mathcal{C}_r), h, r, p, s)$, $\mathcal{C}_r$ is the consensus
    result at round $r$, $p =$ promoter indicator, $s =$ signature.}
  \end{figure}

\end{frame}

\subsection{Protocol Overview}
\begin{frame}{Protocol Overview}
  \begin{enumerate}
    \item $n$ lucky nodes are selected at random to act as promoters.
    \item Promoters run a BFT (Byzantine Fault Tolerant) consensus algorithm to
      agree on a set of CP blocks.
    \item Disseminate the consensus result (the CP blocks).
    \item Repeat for next round.
    \item Any interested node can validate that their transaction.
  \end{enumerate}
\end{frame}

\subsection{Promoter Registration}
\begin{frame}{Promoter Registration}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-1}
  \centering
  \caption{We start in the state where $\mathcal{C}_{r - 1}$ has just been agreed but
    not yet propagated.}
  \end{figure}

\end{frame}

\begin{frame}{Promoter Registration}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-2}
  \centering
  \caption{Nodes receive consensus result and set promoters indicator $p$, then
    send the new CP blocks to promoters of round $r$.}
  \end{figure}

\end{frame}

\begin{frame}{Promoter Registration}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-3}
  \centering
  \caption{Transactions carry on as usual in round $r$. Note that our CP blocks
    (round $r - 1$) has not reached consensus yet.}
  \end{figure}

\end{frame}

\begin{frame}{Promoter Registration}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-4}
  \centering
  \caption{CP blocks at round $r-1$ should be in $\mathcal{C}_r$. If we're
    lucky: $\texttt{h}(k_i || c_{i,j}) < T$, then we're responsible for
    consensus of round $r + 1$.}
  \end{figure}

\end{frame}

\subsection{Consensus}
\begin{frame}{Consensus}
  \begin{enumerate}
    \item Nodes send CP blocks to the promoters.
    \item The promoters' identities are encoded in the consensus result.
    \item Promoters run the some asynchronous BFT consensus algorithm to agree on
      a set of CP blocks---$\mathcal{C}_r$.
    \item $n \ge 3t + 1$ is the optimal for BFT consensus.
    \item $\mathcal{C}_r$ and the signatures are disseminated.
    \item Nodes create new CP blocks when they receive $t + 1$ good signatures and
      $\mathcal{C}_r$.
  \end{enumerate}

  \begin{theorem}
    Assuming the promoters satisfy $n \ge 3t + 1$. The promoter registration and
    the consensus protocol satisfies agreement, total order and liveness.
  \end{theorem}
\end{frame}

\subsection{Validation}

\begin{frame}{Validation}
Assume node $u$ is aware of all the past consensus results $\mathcal{C}_r$.
Suppose $u$ wish to validate $t_{i,j}$. It performs the following.

\begin{enumerate}
\item Determine the pair $t_{i', j'}$.
\item Find the agreed enclosure for $t_{i,j}$ and $t_{i', j'}$ from
  $\mathcal{C}_r$, otherwise return ``unknown''.
%\item Find the smallest $c_{i', b}$ such that $b > j'$ is in consensus.
\item Query $i$ and $i'$ for the agreed pieces and ensure hash pointers are
  correct. Otherwise return ``unknown''.
\item Check that $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are
  created correctly using \texttt{newtx}. Otherwise return ``invalid''.
\item Check the checkpoints $c_{i, k}$ and $c_{i', k'}$ that immediately follow
  $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are created in the
  same round, i.e. $\texttt{round}(c_{i, k}) = \texttt{round}(c_{i', k'})$.
  Otherwise return ``invalid''.
\item Return ``valid''.
\end{enumerate}
\end{frame}

\begin{frame}{Validation}
  In essence, given a TX, ask the receiver to proof that it has a set of
  transactions that produces some CP block, the CP block should be in the
  consensus result and the pair of the TX should be in that set.

  \begin{theorem}
    If at least one party of every transaction is honest, then forking and other
    forms of tampering is guaranteed to be detected if the malicious party is
    alive.
  \end{theorem}

  \begin{block}{Question}
    Given a starting digest and an ending digest, is it possible for the prover
    to proof to the verifier that it has a set of blocks linked by hash pointers
    that ``fits'' inside those digests in \emph{zero knowledge}.
  \end{block}
\end{frame}

\subsection{Implementation and Experimentation Results}
\begin{frame}{Implementation (4 weeks ago)}
  \begin{itemize}
    \item Currently on going---using Python and Twisted
    \item Completed BFT consensus algorithm
    \item Completed local TrustChain with Checkpoints
    \item Next step is networked TrustChain and validation protocol
  \end{itemize}
\end{frame}

\begin{frame}{Implementation}
  \begin{itemize}
    \item Completed BFT consensus algorithm, with erasure coding
    \item Completed TrustChain with Checkpoints, with experiments on DAS5
    \item Validation function is working
    \item Next step is to build validation into a network protocol and do experiments
  \end{itemize}
\end{frame}

\begin{frame}{Consensus duration vs population size (static neighbour)}
  \begin{figure}[h]
  \includegraphics[width=0.95\textwidth]{result-normal-cons-vs-pop}
  \centering
  \end{figure}
\end{frame}

\begin{frame}{Throughput vs population size (static neighbour)}
  \begin{figure}[h]
  \includegraphics[width=0.95\textwidth]{result-normal-throughput-vs-pop}
  \centering
  \end{figure}
\end{frame}

\begin{frame}{Consensus duration vs population size (random neighbour)}
  \begin{figure}[h]
  \includegraphics[width=0.95\textwidth]{result-rand-cons-vs-pop}
  \centering
  \end{figure}
\end{frame}

\begin{frame}{Throughput vs population size (random neighbour)}
  \begin{figure}[h]
  \includegraphics[width=0.95\textwidth]{result-rand-throughput-vs-pop}
  \centering
  \end{figure}
\end{frame}

>>>>

\usepackage{amsfonts}
\usepackage{graphicx}
\graphicspath{{../figures/}}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}

\newcommand{\A}{\mathcal{A}}
\newcommand{\C}{\mathcal{C}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\E}{\mathcal{E}}

\algnewcommand\algorithmicupon{\textbf{Upon}}
\algnewcommand\Upon{\item[\algorithmicupon]}

\algnewcommand\algorithmicsend{\textbf{Send}}
\algnewcommand\Send{\item[\algorithmicsend]}

\title{Blockchain Consensus Protocol with Horizontal Scalability}

% A subtitle is optional and this may be deleted
% \subtitle{Optional Subtitle}

@@

% \date{Conference Name, 2013}
\date{\today}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

@@

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
\AtBeginSubsection[] % Do nothing for \subsection*
{
  % do nothing
}

\AtBeginSection[] % Do nothing for \section*
{
  \begin{frame}<beamer>
  \frametitle{Outline}
  \tableofcontents[currentsection]
  \end{frame}
}

@@

% Section and subsections will appear in the presentation overview
% and table of contents.
\section{Introduction}
\subsection{Motivation}
\begin{frame}{\subsecname}{}
  \begin{itemize}
    \item Blockchain systems offer an alternative to central authorities for the first time
    \item Market cap and trade volume figures indicate they are here to stay
    \item Early blockchain systems are not scalable (7 TX/s for Bitcoin)
    \item Parameter tuning leads to centralisation
  \end{itemize}
\end{frame}

\subsection{Research question}
\begin{frame}{\subsecname}
    How do we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
    \emph{scalable} and can reach \emph{global consensus?}
\end{frame}

\subsection{Intuition}
\begin{frame}{\subsecname}{}
  \begin{itemize}
    \item A restaurant owner does not report all of its transactions with an central authority
    \item Ocassionally a customer may leave without paying and this event is reported to a central authority
    \item Our blockchain system achieves scalability using the same idea
  \end{itemize}
\end{frame}

\section{System architecture}
\subsection{System model}
\begin{frame}{\subsecname}
  \begin{itemize}
    \item Population size is $N$
    \item $n$ nodes are facilitators, $t$ nodes are malicious (Byzantine)
    \item $n \ge 3t + 1$
    \item $N \ge n + t$
    \item Purely asynchronous channels with eventual delivery
    \item Public key infrastructure
    \item Random oracle model
    \item Application neutural
  \end{itemize}
\end{frame}

\subsection{Architecture overview}
\begin{frame}{\subsecname}
  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{architecture}
  \centering
  \end{figure}
\end{frame}

\subsection{Extended TrustChain}
\begin{frame}{\subsecname}
  \begin{itemize}
    \item Everyone has their own chain
    \item Two types of blocks, transaction (TX) blocks and checkpoint (CP) blocks
    \item A transaction involves two parties and results in two TX blocks
    \item A CP block captures the chain state
    \item TX and CP blocks are chained together using hash pointers
  \end{itemize}
\end{frame}

\begin{frame}{\subsecname}
  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-bad-cp}
  \centering
  \caption{TX block is a six-tuple: $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), i, txid, pk_v, m, sig_u \rangle$.
    CP block is a five-tuple: $c_{u, i} = \langle \textsf{H}(b_{u, i-1}), i, \textsf{H}(\C_r), r, sig_u \rangle$.}
  \end{figure}
\end{frame}

\subsection{Consensus protocol}
\begin{frame}{\subsecname overview}
  \begin{enumerate}
    \item In round $r$, $n$ out of $N$ lucky nodes are selected at random to act as facilitators
    \item Facilitators run a BFT (Byzantine Fault Tolerant) consensus algorithm to
      agree on a set of CP blocks
    \item Disseminate the consensus result $\C_r$, i.e. the CP blocks
    \item Repeat
  \end{enumerate}
\end{frame}

\begin{frame}{\subsecname properties}
\label{def:consensus}
$\forall r \in \mathbb{N}$, the following properties must hold.
\begin{itemize}
    \item \emph{Agreement}:
        If one correct node outputs a list of facilitators $\F_r$,
        then every node outputs $\F_r$
    \item \emph{Validity}:
        If any correct node outputs $\F_r$, then 
            \begin{enumerate}
                \item $|\C_r| \ge N - t$ must hold for the $\C_r$ which was used to create $\F_r$,
                \item $\F_r$ must contain at least $n - t$ honest nodes and
                \item $|\F_r| = n$.
            \end{enumerate}
    \item \emph{Fairness}:
        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
    \item \emph{Termination}:
        Every correct node eventually outputs some $\F_r$.
\end{itemize}
\end{frame}

\begin{frame}{\subsecname}
  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-1}
  \centering
  \caption{Suppose we are in a state where $\mathcal{C}_{r - 1}$ has just been agreed by some facilitators but not yet propagated.}
  \end{figure}
\end{frame}

\begin{frame}{\subsecname}
  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-2}
  \centering
  \caption{Nodes receive consensus result $\mathcal{C}_{r - 1}$,
    compute the new facilitators $\mathcal{F}_{r-1}$,
    and send out the new CP blocks.}
  \end{figure}
\end{frame}

\begin{frame}{\subsecname}
  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-3}
  \centering
  \caption{Transactions carry on as usual in round $r$,
  while facilitators are trying to reach consensus on the new CP blocks concurrently.}
  \end{figure}
\end{frame}

\begin{frame}{\subsecname}
  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-4}
  \centering
  \caption{CP blocks at round $r-1$ should be in $\mathcal{C}_r$.
  The first $n$ nodes ordered by $\textsf{H}(\mathcal{C}_r || pk)$ become $\mathcal{F}_r$.}
  \end{figure}
\end{frame}

\subsection{Transaction protocol}
\begin{frame}{\subsecname}
\begin{itemize}
\item Request (\texttt{tx\_req}) and response (\texttt{tx\_resp}) protocol
\item Two TX blocks containing the same $txid$ are generated
\item Non-blocking
\end{itemize}
\end{frame}

\begin{frame}{\subsecname}
Create $t_{u, h}$ and send $\langle \texttt{tx\_req}, t_{u, h} \rangle$ to start a transaction.
\vspace{5 mm}
\begin{algorithmic}
    \Upon $\langle \texttt{tx\_req}, t_{v, j} \rangle$ from $v$
    \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{v, j}$ \Comment unpack $t_{v, j}$
    \State $\textsf{new\_tx}(pk_u, m, txid)$ \Comment create and store $t_{u, i}$
    \State store $t_{v, j}$ as the pair of $t_{u, h}$
    \State send $\langle \texttt{tx\_resp}, t_{u, h} \rangle$ to $v$

    \Upon $\langle \texttt{tx\_resp}, t_{v, j} \rangle$ from $v$  
    \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{v, j}$ \Comment unpack $t_{v, j}$
    \State store $t_{v, j}$ as the pair of the TX with identifier $txid$
\end{algorithmic}
\end{frame}

\subsection{Validation protocol}
\begin{frame}{\subsecname overview}
\begin{itemize}
\item Request (\texttt{vd\_req}) and response (\texttt{vd\_resp}) protocol
\item Transactions are in three states---\emph{valid}, \emph{invalid} and \emph{unknown}
\item The goal is to identify which state a given transaction is in
\end{itemize}
\end{frame}

\begin{frame}{\subsecname properties}
\begin{itemize}
    \item \emph{Correctness}:
        The validation protocol outputs the correct result
        according to the aforementioned validity definition.
    \item \emph{Agreement}:
        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
    \item \emph{Liveness}:
        Any valid transactions can be validated eventually.
\end{itemize}
\end{frame}

\begin{frame}{Validity definition}
\textbf{Function} $\textsf{get\_validity}()$ validates the transaction $t_{u, i}$
\vspace{5 mm}
\begin{algorithmic}
    \State Check that $v$ sent the correct agreed fragment,
    otherwise return \emph{unknown}.

    \State
    \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{u, i}$
    \If{number of blocks of $txid$ in $F_{v, j} \ne 1$}
        \State \Return \emph{invalid}
    \EndIf
    \State \Comment TX exists

    \State $\langle \_, \_, txid', pk'_u, m', \_ \rangle \gets t_{v, j}$
    \If{$m \ne m' \vee pk_u \ne pk'_u$}
        \State \Return \emph{invalid}
    \EndIf
    \State \Comment no tampering
    \State \Return \emph{valid}
\end{algorithmic}
\end{frame}

\begin{frame}{\subsecname}
Send $\langle \texttt{vd\_req}, txid \rangle$ to $v$ to begin.
\vspace{5 mm}
\begin{algorithmic}
    \Upon $\langle \texttt{vd\_req}, txid \rangle$ from $v$
        \State $t_{u, i} \gets \text{the transaction identified by } txid$
        \State $F_{u, i} \gets \textsf{agreed\_fragment}(t_{u, i})$
        \State send $\langle \texttt{vd\_resp}, txid, F_{u, i} \rangle$ to $v$

    \Upon $\langle \texttt{vd\_resp}, txid, F_{v, j} \rangle$ from $v$
        \State $t_{u, i} \gets \text{the transaction identified by } txid$
        \State set the validity of $t_{u, i}$ to $\textsf{get\_validity}(t_{u, i}, F_{v, j})$
\end{algorithmic}
\end{frame}

\section{Analysis of correctness and performance}

\subsection{Correctness of the consensus protocol}
\begin{frame}{\subsecname}
\begin{theorem}
For all rounds,
the consensus protocol satisfies agreement, validity, fairness and termination.
\end{theorem}
\begin{proof}(sketch)
Because consensus result are eventually delivered and the properties of ACS,
we get agreement, validity and termination.
Fairness is from the fact that we model $\textsf{H}(\cdot)$ as a random oracle (RO) and the input to the RO is different for every node,
thus the list of nodes ordered by $\textsf{H}(\C_r || pk )$ is a random permutation of those nodes.
\end{proof}
\end{frame}

\subsection{Correctness of the validation protocol}
\begin{frame}{\subsecname}
\begin{theorem}
If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
\end{theorem}
\begin{proof}(sketch)
Proof by contradiction.
For this attack to work, the adversary must be able to create two different fragments but with the same checkpoint enclosure.
We model $\textsf{H}(\cdot)$ as a RO, so the adversary need to query the RO a exponential\footnote{In terms of the security parameter.} number of times.
But the adversary can only query the RO a polynomial number of times.
\end{proof}
\end{frame}

\subsection{Linear global throughput argument}
\begin{frame}{\subsecname}
\begin{itemize}
  \item Throughput has a notion of time but our model is asynchronous
  \item Additional assumptions needed to make the argument---every unit of communication takes a non-negligible amount of time to process
  \item Bandwidth relation---$NC \ge r_{\text{tx}} l$, where $l$ is $O(N)$
  \item If $r_\text{tx}$ satisfies the inequality, then LHS and RHS grows at the same rate,
  thus we have linear global throughput
\end{itemize}
\end{frame}

\section{Experimental results}
\begin{frame}{Implementation and experiment setup}
  \begin{itemize}
    \item Prototype implementation on Github\footnote{\url{https://github.com/kc1212/consensus-thesis-code}}
    \item SHA256 for hash functions and Ed25519 for digital signature
    \item Experiment on the DAS-5\footnote{\url{http://www.cs.vu.nl/das5/}}
  \end{itemize}
\end{frame}

\begin{frame}{Throughput vs population size (random neighbour)}
  \begin{figure}[h]
  \includegraphics[width=1.0\textwidth]{neighbour-random/throughput-vs-population}
  \centering
  \end{figure}
\end{frame}

\begin{frame}{Throughput vs population size (fixed neighbour)}
  \begin{figure}[h]
  \includegraphics[width=1.0\textwidth]{neighbour-fixed/throughput-vs-population}
  \centering
  \end{figure}
\end{frame}

\section{Conclusion}
\begin{frame}{\secname}
    How do we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
    \emph{scalable} and can reach \emph{global consensus?}
\begin{itemize}
  \item[\checkmark] Fault tolerant up to $t$ nodes
  \item[\checkmark] Horizontal scalability
  \item[\checkmark] Global consensus
\end{itemize}
\end{frame}

\begin{frame}{Future work}
\begin{itemize}
\item Improve fault tolerance
\item Improve fork detection
\item Analyse the system in the permissionless environment
\item Concrete application
\end{itemize}
\end{frame}
